Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-17 concentration.tex
updated sections 5 and 6 with respect to changes in section 3

\section{Concentration for $c_{\mathcal{P}, n}(k)$ (Proving Proposition \ref{prop:concentration_local_clustering_P_n})}
\label{sec:concentration_c_P_n}

@@

Recall the adjusted triangle count function \eqref{eq:def_tilde_triangle_indicator}
\[
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal_n}.
\]
and the concentration set
\[
	\Kcal_C(k) = \left\{p \in \R : \frac{k - C \sqrt{k \log(k)}}{\xi_{\alpha,\nu}} \le e^{\frac{y}{2}}
		\le \frac{k + C \sqrt{k \log(k)}}{\xi_{\alpha,\nu}} \right\},
\]
We will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\Pcal,n}(k,C) = \sum_{p \in \Pcal_n \cap K_C(k)} \ind{D_{\Pcal,n}(p) = k} \widetilde{T}_{\Pcal,n}(p),
\end{equation}
where
\[
	\widetilde{T}_{\Pcal,n}(p) := \sum_{(p_1, p_2) \in 2^{\Pcal_n}} \widetilde{T}_{\Pcal,n}(p,p_1,p_2).
\]

By Lemma \ref{lem:clustering_error_T_term} and a concentration argument it follows that for $k_n \to \infty$
\[
	\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)} = (1+\smallO{1})\Exp{T_{\Pcal}(k_n)},
\]
for a appropriately selected $C > 0$. We conclude that the main contribution of triangles of degree $k_n$ is given by $\widetilde{T}_{\Pcal,n}(k_n,C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} we need to show that $\widetilde{T}_{\Pcal,n}(k_n,C)$ is sufficiently concentrated around its mean.

\subsection{Proving Proposition \ref{prop:concentration_local_clustering_P_n}}

We start with the concentration result for $\widetilde{T}_{\Pcal,n}(k_n,C)$.

\begin{proposition}[Concentration $\widetilde{T}_{\Pcal,n}(k_n,C)$]\label{prop:concentration_tilde_T_P_n}
Let $\alpha > \frac{1}{2}$, $0 < \varepsilon < \min\{2\alpha - 1,1\}$ and let $(k_n)_{n \ge 1}$ be any increasing sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\widetilde{T}_{\Pcal,n}(k_n, C)^2} = \left(1 + \smallO{1}\right)\Exp{\widetilde{T}_{\Pcal,n}(k_n, C)}^2.
\]
\end{proposition}

The proof of this proposition is lengthy and we therefore postpone it till Section \ref{ssec:concentration_tilde_T}. The remaining of this section will be devoted to prove Proposition \ref{prop:concentration_local_clustering_P_n}. We first show that $\Exp{c_{\Pcal,n}(k_n)} = (1+o(1))\Exp{c^\ast_{\Pcal,n}(k_n)}$.

\begin{lemma}\label{lem:L1_convergence_c_Pcal_n}
Let $\nu > 0$, $\alpha > \frac{1}{2}$ and $k_n = o\left(n^{\frac{1}{2\alpha + 1}}\right)$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
\]
\end{lemma}

\begin{proof}
Let $0 < \delta < 1$ and define the following two events
\begin{align*}
	A_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le \Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}\right\}\\
	B_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le n\right\},
\end{align*}
so that
\[
	1 = \ind{A_n} + \ind{A_n^c}\ind{B_n} + \ind{B_n^c}.
\]
Next we note that on the event $A_n$
\[
	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
	\le \frac{\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}{\Exp{N_{\Pcal,n}(k_n)}+\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}
	\le \Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}},
\]
and on the event $B_n$
\[
	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
	\le \frac{n}{\Exp{N_{\Pcal,n}(k_n)} + n} \le 1.
\]
Therefore we have
\begin{align*}
	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|}
	&= \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|}\\
	&\le \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n}}\Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}} \\
	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} \\
	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}}
\end{align*}
Since $\Exp{N_{\Pcal,n}(k_n)} \to \infty$, the first term is clearly $\smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}$. For the third term we have
\begin{align*}
	\Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}} 
	&= \bigO{\Prob{B_n^c}}
		= \bigO{\Prob{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| > n}}\\
	&= \bigO{e^{-\frac{n^2}{\Exp{N_{k_n}} + n}}} = \bigO{e^{-n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
\end{align*}
Hence we are left to show that
\begin{equation}\label{eq:L1_convergence_second_term}
	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
\end{equation}
By writing
\begin{align*}
	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \frac{T_{\Pcal,n}(k_n,\varepsilon)- \widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
\end{align*}
and using Lemma~\ref{lem:clustering_error_T_term} we get
\begin{align*}
	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}}
	&\le \Exp{\frac{(1+\smallO{1})\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
		\ind{A_n^c}} + \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
\end{align*}
For the last step we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to get
\begin{align*}
	\Exp{\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\ind{A_n^c}}
	&\le \Exp{\left(\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}
		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\right)^2}^{\frac{1}{2}} \Prob{A_n^c}^{\frac{1}{2}}\\
	&= \left(1 + \smallO{1}\right) 	
		\frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
		\Prob{A_n^c}^{\frac{1}{2}}\\
	&= \bigO{\Exp{c_{\Pcal, n}^\ast(k_n)}}\Prob{A_n^c}^{\frac{1}{2}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}},
\end{align*}
since $\Prob{A_n} = o(1)$. This establishes \eqref{eq:L1_convergence_second_term} and hence we conclude that
\[
	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
\]
\end{proof}

We are now in shape to prove the main result of this section.

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
Again, we write
\begin{align*}
	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \frac{\left(T_{\Pcal,n}(k_n)-\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)\right)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}},
\end{align*}
so that by Lemma \ref{lem:clustering_error_T_term},
\[
	\Exp{c_{\Pcal,n}^\ast(k_n)} = \frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
\]
Therefore, by Lemma \ref{lem:L1_convergence_c_Pcal_n},
\begin{align*}
	\Exp{\left|c_{\Pcal,n}(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
	&\le \Exp{\left|c_{\Pcal,n}^\ast(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
		+ \Exp{\left|c_{\Pcal,n}(k_n) - c_{\Pcal,n}^\ast(k_n)\right|}\\
	&= \frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}} + \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
		\numberthis \label{eq:concentration_local_clustering_P_n_term1}
\end{align*}
Next, we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)^2} 
		- \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	&= \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}},
\end{align*}
which together with~\eqref{eq:concentration_local_clustering_P_n_term1} finishes the proof.
\end{proof}

>>>>

\section{Concentration for $c_{\Pcal, n}(k)$ (Proving Proposition \ref{prop:concentration_local_clustering_P_n})}
\label{sec:concentration_c_P_n}

In this section we establish a concentration result for the local clustering function $c_{\Pcal, n}^\ast(k)$ in the finite model $G_{\Pcal,n}$. Similar to the previous section we will focus on those nodes $p = (x,y) \in \Kcal_{C}(k_n)$.

@@

Recall the adjusted triangle count function \eqref{eq:def_tilde_triangle_indicator}
\[
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal_n}.
\]
and the concentration set
\[
	\Kcal_C(k_n) = \left\{p \in \R : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\]
with
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\]
We will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\Pcal,n}(k, C) = \sum_{p \in \Pcal_n \cap K_C(k)} \ind{D_{\Pcal,n}(p) = k} \widetilde{T}_{\Pcal,n}(p),
\end{equation}
where
\[
	\widetilde{T}_{\Pcal,n}(p) := \sum_{(p_1, p_2) \in \Pcal\setminus (0,y)}^{\ne} \widetilde{T}_{\Pcal,n}(p,p_1,p_2),
\]
where the sum is over all distinct pairs $(p_1, p_2) \in \Pcal \setminus \{(0,y)\}$.

Next we note that the second result from Corollary~\ref{cor:approximation_triangle_integral} states that
\begin{equation}\label{eq:asymptotic_equivalence_contraint_triangle_count}
	\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)} = (1+\smallO{1})\Exp{T_{\Pcal}(k_n)},
\end{equation}
so that we can conclude that the main contribution of triangles of degree $k_n$ is given by $\widetilde{T}_{\Pcal,n}(k_n, C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} it suffices to show that $\widetilde{T}_{\Pcal,n}(k_n,C)$ is sufficiently concentrated around its mean. This is done in the following proposition.

%\subsection{The proof of Proposition \ref{prop:concentration_local_clustering_P_n}}
%
%We start with the concentration result for $\widetilde{T}_{\Pcal,n}(k_n)$.

\begin{proposition}[Concentration $\widetilde{T}_{\Pcal,n}(k_n,C)$]\label{prop:concentration_tilde_T_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then for any $C > 0$, as $n \to \infty$,
\[
	\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} = \left(1 + \smallO{1}\right)\Exp{\widetilde{T}_{\Pcal,n}(k_n, C)}^2.
\]
\end{proposition}

We postpone the proof of this proposition postpone till Section \ref{ssec:concentration_tilde_T} and first use it to prove Proposition \ref{prop:concentration_local_clustering_P_n}.

%\begin{lemma}\label{lem:L1_convergence_c_Pcal_n}
%Let $\nu > 0$, $\alpha > \frac{1}{2}$ and $k_n = o\left(n^{\frac{1}{2\alpha + 1}}\right)$. Then, as $n \to \infty$,
%\[
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{lemma}
%
%\begin{proof}
%Let $0 < \delta < 1$ and define the following two events
%\begin{align*}
%	A_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le \Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}\right\}\\
%	B_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le n\right\},
%\end{align*}
%so that
%\[
%	1 = \ind{A_n} + \ind{A_n^c}\ind{B_n} + \ind{B_n^c}.
%\]
%Next we note that on the event $A_n$
%\[
%	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
%	\le \frac{\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}{\Exp{N_{\Pcal,n}(k_n)}+\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}
%	\le \Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}},
%\]
%and on the event $B_n$
%\[
%	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
%	\le \frac{n}{\Exp{N_{\Pcal,n}(k_n)} + n} \le 1.
%\]
%Therefore we have
%\begin{align*}
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|}
%	&= \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|}\\
%	&\le \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n}}\Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}} \\
%	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} \\
%	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}}
%\end{align*}
%Since $\Exp{N_{\Pcal,n}(k_n)} \to \infty$, the first term is clearly $\smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}$. For the third term we have
%\begin{align*}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}} 
%	&= \bigO{\Prob{B_n^c}}
%		= \bigO{\Prob{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| > n}}\\
%	&= \bigO{e^{-\frac{n^2}{\Exp{N_{k_n}} + n}}} = \bigO{e^{-n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{align*}
%Hence we are left to show that
%\begin{equation}\label{eq:L1_convergence_second_term}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{equation}
%By writing
%\begin{align*}
%	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%	+ \frac{T_{\Pcal,n}(k_n,\varepsilon)- \widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%\end{align*}
%and using Lemma~\ref{lem:clustering_error_T_term} we get
%\begin{align*}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}}
%	&\le \Exp{\frac{(1+\smallO{1})\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%		\ind{A_n^c}} + \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{align*}
%For the last step we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to get
%\begin{align*}
%	\Exp{\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\ind{A_n^c}}
%	&\le \Exp{\left(\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}
%		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\right)^2}^{\frac{1}{2}} \Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \left(1 + \smallO{1}\right) 	
%		\frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%		\Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \bigO{\Exp{c_{\Pcal, n}^\ast(k_n)}}\Prob{A_n^c}^{\frac{1}{2}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}},
%\end{align*}
%since $\Prob{A_n} = o(1)$. This establishes \eqref{eq:L1_convergence_second_term} and hence we conclude that
%\[
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{proof}

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
Again, we write
\begin{align*}
	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \frac{\left(T_{\Pcal,n}(k_n)-\widetilde{T}_{\Pcal,n}(k_n,C)\right)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}},
\end{align*}
so that by~\eqref{eq:asymptotic_equivalence_contraint_triangle_count},
\begin{equation}\label{eq:concentration_local_clustering_P_n_main}
	\Exp{c_{\Pcal,n}^\ast(k_n)} = \frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
\end{equation}
%Therefore, by Lemma \ref{lem:L1_convergence_c_Pcal_n},
%\begin{align*}
%	\Exp{\left|c_{\Pcal,n}(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
%	&\le \Exp{\left|c_{\Pcal,n}^\ast(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
%		+ \Exp{\left|c_{\Pcal,n}(k_n) - c_{\Pcal,n}^\ast(k_n)\right|}\\
%	&= \frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
%		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}} + \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
%		\numberthis \label{eq:concentration_local_clustering_P_n_term1}
%\end{align*}
Next, we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	&= \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}},
\end{align*}
which together with~\eqref{eq:concentration_local_clustering_P_n_main} yields the required result.
\end{proof}

\subsection{Joint neighborhoods and degrees in $G_{\Pcal,n}(\alpha,\nu)$}

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $G_{\Pcal,n}$ and subsequently the joint neighborhoods of two points $p, p^\prime \in \Rcal_n$. We perform the analysis in this section. The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma ?? and Lemma ??.

\subsubsection*{Neighborhoods}

We start with analyzing joint neighborhoods in $G_{\Pcal,n}$. Let $p, p^\prime \in \Rcal_n$. Then we denote by $\Ncal_{\Pcal,n}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $G_{\Pcal,n}$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by $\Ncal_{\Pcal,n}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish lower bounds on the expect number of these disjoint neighbors $\Exp{\Ncal_{\Pcal,n}(p\Delta p^\prime)}$ as well as an asymptotic expression for the expected number of joint neighbors. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.

\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
    \draw node at (-0.2,0.7) {$p$};

    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (1,0.5) {};
    \draw node at (0.8,0.2) {\color{blue}$p^\prime$};

	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
    \draw[domain=2.2840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-1)-0.5});
    \draw[domain=-0.2840:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(1-\x)-0.5});

    \draw[dashed,thick,black,line width=1pt] (-6,2.5835) -- (6,2.5835);

    \draw node at (-7,2.5835) {$h(y)$};
    \draw node at (-7,3.3918) {\color{blue}$h_1(p^\prime)$};
    \draw node at (7,2.7189) {\color{blue}$h_2(p^\prime)$};

    \draw node at (-4,3.5) {\color{blue}$x_1 = x^\prime - e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (4.5,3) {\color{blue}$x_1 = x^\prime + e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (-4.5,1) {$x_1 = x - e^{\frac{y + y_1}{2}}$};
    \draw node at (2,1.5) {$x_1 = x + e^{\frac{y + y_1}{2}}$};

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
    \draw node at (-0.2,0.7) {$p$};

    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (2.5,0.5) {};
    \draw node at (2.3,0.2) {\color{blue}$p^\prime$};

	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
    \draw[domain=3.7840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-2.5)-0.5});
    \draw[domain=1.2160:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(2.5-\x)-0.5});
    \draw[domain=0:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x+9.5)-0.5});

    \draw[dashed,black,line width=1pt] (-6,2.0055) -- (6,2.0055);

%    \draw node at (-7,2.5835) {$h(y)$};
    \draw node at (-7,3.7801) {\color{blue}$h_1(p^\prime)$};
    \draw node at (7,2.0055) {\color{blue}$h_2(p^\prime)$};

%    \draw[dotted,thick,black] (4.5208,2.0174) -- (4.5208,0);
%    \draw[dotted,thick,black] (4.5208,2.0174) -- (6,2.0174);

%    \draw node at (4.5208,-0.5) {$w_x(p,p^\prime)$};
%    \draw node at (7,2.0174) {$w_y(p,p^\prime)$};

    \draw node at (0,3.5) {\color{blue}$x_1 = x^\prime + e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (-2,1.6) {\color{blue}$x_1 = x^\prime - e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (-4.5,1) {$x_1 = x - e^{\frac{y + y_1}{2}}$};
    \draw node at (2,1.5) {$x_1 = x + e^{\frac{y + y_1}{2}}$};

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ when $e^{\frac{y + y^\prime}{2}} < |x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n_2}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
    \draw node at (-0.2,1.3) {$p$};

    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (4.5,0.5) {};
    \draw node at (4.3,0.2) {\color{blue}$p^\prime$};

	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
    \draw[domain=5.7840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-4.5)-0.5});
    \draw[domain=3.2160:-5,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(4.5-\x)-0.5});
    \draw[domain=2:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x+7.5)-0.5});

    \draw[dashed,black,line width=1pt] (-6,0.8563) -- (6,0.8563);

    \draw node at (-7,0.8563) {$h(p,p^\prime)$};
    \draw node at (7,0.3109) {\color{blue}$h_1(p^\prime)$};

    \draw [decorate,decoration={brace},line width=1pt] (3.2160,-0.1) -- (1.6487,-0.1);
    \draw node at (2.5,-0.45) {$d(p,p^\prime)$};

    \draw [dotted,line width=1pt] (2.5298,0.8563) -- (2.5298,1.7);
    \draw node at (2.5298,2) {$x^\ast(p,p^\prime)$};

%    \draw[dotted,thick,black] (4.5208,2.0174) -- (4.5208,0);
%    \draw[dotted,thick,black] (4.5208,2.0174) -- (6,2.0174);
